---
output: 
  rmarkdown::github_document:
    toc: true
    toc_depth: 2
editor_options: 
  chunk_output_type: console
---
# Bump

## Abstract

Analysis of late night shows looking at the bump in google trends hits following a show. The Colbert Report discussed the "Colbert-bump", where a guest or topic of the show received a boost in attention following the show. This repository and analysis set out to quantify and summarize the gain in attention across various late night shows.

```{r}
tictoc::tic("Compile report")
suppressPackageStartupMessages({
  library(dplyr)
  library(ggplot2)
  library(ggridges)
})
knitr::opts_chunk$set(warning = FALSE, message = FALSE, cache = FALSE)
```

## Gathering the data

Before we can search the show we need a list of episodes, topics, and guests to work with. To create such a list I had co-pilot aid me in creating an `rvest` script to search Wikipedia articles for 4 different late night shows: The Daily Show (various hosts), Last Week Tonight (John Oliver), Late Night (Seth Meyers), and The Late Show (Stephen Colbert). Program information contained topics, guests, musical/entertainment guests, and occasionally a description.

```{r}
load("output/shows2.rda")
shows2 %>% str()

shows2 %>%
  count(Show, topic = !is.na(Topic)) %>%
  filter(topic == TRUE)
```

After a variable to compare is selected, `gtrendsR` is used to look up the relative google hits for each simplified term. A period of 14 days on either side of the episode air date was retrieved from google trends. Google trends was available since Jan 1, 2004. I define the following measure to quantify the gain in search hits around the air date of the episode.

```
ratio = sum of hits in the previous 7 days (minimum of 1) / sum of hits in the proceeding 7 days
```

## Topic

Let's consider Topic data for Last Week Tonight and The Daily show.

```{r}
load("output/gtrend_topic.rda")
gtrend_topic <- gtrend_topic %>%
  select(-c(Topic, Topic_simplified, Topic_simplified_split))
gtrend_topic %>% str()
```

Not everything returned the expected google trends data. Let's review how that broke down.

```{r}
## Count Show-episode-search worked
gtrend_topic %>%
  count(Show, Host, Season, Episode, keyword) %>%
  arrange(-n) %>%
  count(Show, Host, worked = n >= 7)

## Topics that Worked
(worked_topic <- gtrend_topic %>%
    count(Show, Season, Episode, keyword) %>%
    filter(n >= 7))
```

Still, of the remaining data we can apply this analysis.

## Show-grained visuals

This illustrates the ratio timeframe, one week beforehand is shaded in orcid purple, while a one week after is shaded in salmon red. The ratio is the after over before (with a minimum of 1). The ratio of the different topics are displayed in parentheses.

```{r}
gdat <- worked_topic %>%
  ## Filter data to those that worked
  inner_join(gtrend_topic, .) %>%
  ## Create measures
  mutate(days_relative_to_Air_Date = (as.Date(date) - Air_Date) %>% as.numeric()) %>%
  group_by(Show, Host, Season, Episode, Air_Date, keyword) %>%
  summarize(
    sum_hits_before = max(sum(hits[days_relative_to_Air_Date < 0 & days_relative_to_Air_Date >= -7], na.rm = TRUE), 1),
    sum_hits_after = sum(hits[days_relative_to_Air_Date > 0 & days_relative_to_Air_Date <= 7], na.rm = TRUE),
    ratio = sum_hits_after / sum_hits_before
  ) %>%
  ungroup() %>%
  ## Join back to data.
  inner_join(., gtrend_topic %>%
               select(Show, Host, Season, Episode,
                      Air_Date, keyword, date, hits)) %>%
  mutate(keyword2 = paste0(keyword, " (", (ratio %>% round(2)), ")"),
         start_purple = Air_Date - 7,
         end_purple = Air_Date,
         start_red = Air_Date,
         end_red = Air_Date + 7
  ) %>%
  arrange(-ratio)

## Plot
gdat %>%
  head(16 * 29) %>%
  filter(!is.na(date), !is.na(keyword2)) %>%
  mutate(date = as.Date(date)) %>%
  ggplot(aes(date, hits, group = keyword2)) +
  geom_rect(
    aes(xmin = start_purple, xmax = end_purple, ymin = -Inf, ymax = Inf),
    fill = "orchid", color = "grey50", alpha = 0.1, inherit.aes = FALSE) +
  geom_rect(
    aes(xmin = start_red, xmax = end_red, ymin = -Inf, ymax = Inf),
    fill = "salmon", color = "grey50", alpha = 0.1, inherit.aes = FALSE) +
  geom_line() +
  facet_wrap(~keyword2, scales = "free_x") +
  labs(x = "Date", y = "Relative Hits") +
  theme_bw()
```

## Show summaries

Summary of show:

```{r}
## Summary of show
gdat %>%
  group_by(Show, Host) %>%
  summarize(mean_ratio = mean(ratio, na.rm = TRUE),
            median_ratio = median(ratio, na.rm = TRUE),
            sd_ratio = sd(ratio, na.rm = TRUE),
            mean_keywords_per_episode = mean((n() / 29) %>% ceiling(), na.rm = TRUE) / n_distinct(Episode)) %>%
  left_join(., gtrend_topic %>%
              group_by(Show, Host) %>%
              summarize(pct_error = 100 * sum(substring(error, 1, 5) == "Error", na.rm = TRUE) / n())
  )
```

Visual of ratio, log10

```{r}
gdat %>%
  #filter(ratio > 1) %>%
  count(Show, Host, keyword, ratio) %>%
  inner_join(., gdat %>% count(Show, Host) %>% rename(n_keywords = n)) %>%
  inner_join(., gdat %>%
               count(Show, Host, keyword) %>%
               mutate(n_worked = n > 7) %>%
               count(Show, Host, n_worked) %>%
               select(-n_worked) %>%
               rename(n_worked = n)) %>%
  mutate(Show_Host = paste0(Show, ", ", Host, " (",n_worked, " / ", n_keywords, ")")) %>%
  ggplot(aes(x = log10(ratio + .1),
             y = reorder(Show_Host, n_keywords), fill = Show_Host)) +
  geom_density_ridges(scale = .9) + # Adjust the scale to reduce overlap
  theme_bw() +
  theme(legend.position = "none") +
  labs(x = "Ratio [log10]",
       y = "Show, Host (Keywords with valid gtrends / Count of keywords)")
```

Highest ratios:

```{r}
## Keyword by descending ratio
gdat %>%
  filter(ratio != Inf) %>%
  count(Show, Host, Season, Episode, keyword, ratio) %>%
  arrange(-ratio) %>%
  select(-n)
```

Lowest ratios:

```{r}
## Keyword by ascending ratio
gdat %>%
  filter(ratio != Inf) %>%
  count(Show, Host, Season, Episode, keyword, ratio) %>%
  arrange(ratio) %>%
  select(-n)
```


## Limitations

Google trends only go back to 2004/1/1. Google trends had surprisingly low throughput of returning results even when querying spanned several days to avoid API limits. Some trend response seem unrealistically low, for instance all of John Oliver's last week tonight that worked gave ratios of 0 or 1, but never above that.


# Session info

```{r}
tictoc::toc() ## "Compile report"
sessionInfo()
```

